---
title: "p8105_hw2_hf2379"
author: "Holly Finertie"
date: "9/28/2019"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

# Problem 1: 

## Read and clean data sets

```{r}
library(readxl)

trash_wheel = read_excel("./data/HealthyHarborWaterWheelTotals2018-7-28.xlsx", 1, range = "A2:N338") %>% 
  janitor::clean_names() %>% 
  drop_na(dumpster) %>% 
  mutate(sports_balls = round(sports_balls, digits=0)) %>% 
  mutate(sports_balls = as.integer(sports_balls))

precip_2017 = read_excel("./data/HealthyHarborWaterWheelTotals2018-7-28.xlsx", 4, range = "A2:B14") %>% 
  janitor::clean_names() %>% 
  drop_na(total) %>% 
  mutate(year = 2017)

precip_2018 = read_excel("./data/HealthyHarborWaterWheelTotals2018-7-28.xlsx", 3, range = "A2:B14") %>% 
  janitor::clean_names() %>% 
  drop_na(total) %>% 
  mutate(year = 2018)

precip_combined_2017_2018 = 
  bind_rows(precip_2017, precip_2018) %>% 
  mutate(month = month.name[month])
```

Two data sets have been successfully imported and cleaned: (1) trash_wheel and (2) precip_both_years. 


**trash_wheel**: This data set has `r trash_wheel %>% nrow()` observations and `r trash_wheel %>% ncol()` columns describing the weight and volume of trash pulled out of the Inner Harbor in Baltimore by month from `r trash_wheel %>%  select(year) %>% min()` to `r trash_wheel %>% select(year) %>% max()`. Additional variables describe the type of trash pulled (e.g. plastic bottles, cigarette butts, and sports balls). For example, in 2014 Mr. Trash Wheel pulled out `r trash_wheel %>% filter(year==2014) %>% select(weight_tons) %>% sum()` tons and `r trash_wheel %>% filter(year==2014) %>% select(volume_cubic_yards) %>% sum()` cubic yards of trash. Additionally, in 2017 the median number of sports balss in a dumpster pulled from Mr. Trash Wheel was `r trash_wheel %>% filter(year==2017) %>% mutate(balls=as.numeric(sports_balls)) %>% select(sports_balls) %>% min()`. 

**precip_both_years**: This data set has `r precip_combined_2017_2018 %>% nrow()` observations and `r precip_combined_2017_2018 %>% ncol()` columns describing the total precipation in inches by month from `r min(pull(precip_combined_2017_2018, year))` to `r precip_combined_2017_2018 %>% select(year) %>% max()`. For example, from the table we know that the precipitation in January 2018 was `r precip_combined_2017_2018 %>% filter(year==2018) %>% filter(month=="January") %>% select(total)` inches and total precipation in 2018 was `r precip_combined_2017_2018 %>% filter(year==2018) %>% select(total) %>% sum()`
inches. 


## REMEMBER YOU DIDN'T DO THE MEDIAN


# Problem 2: 

```{r}
pols_month = read_csv("./data/pols-month.csv") %>% 
  janitor::clean_names() %>% 
  separate(mon, into = c("year", "month", "day")) %>% 
  mutate(
      month = as.numeric(month),
      month = month.name[month],
      year = as.numeric(year),
      president = recode(prez_gop, `1` = "gop", `2`="gop", `0`="dem")
      ) %>% 
  select(-day, -prez_gop, -prez_dem)

snp = read_csv("./data/snp.csv") %>% 
  janitor::clean_names() %>% 
  separate(date, into = c("month", "day", "year")) %>% 
  select("year", "month", everything()) %>% 
  mutate(
      month = as.numeric(month),
      year = as.numeric(year)) %>% 
  arrange(year, month) %>% 
  mutate(month = month.name[month])

unemploy = read_csv("./data/unemployment.csv") %>% 
  janitor::clean_names() %>% 
  pivot_longer(jan:dec, names_to = "month", values_to = "unemploy_perc") %>% 
  mutate(
    month=recode(month, "jan"="January", "feb"="February", "mar"="March", "apr"="April", "may"="May", "jun"="June", "jul"="July", "aug"="Aug", "sep"="September", "oct"="October", "nov"="November", "dec"="December"))

snp_pols = 
  left_join(pols_month, snp, by = c("month" = "month", "year" = "year"))

final_table_538 = 
    left_join(snp_pols, unemploy, by = c("month" = "month", "year" = "year"))

```








